---
title: "Corpus Descriptive Analysis"
subtitle: "Multi-Model Foundry Subject Rearing Project"
author: "Generated from pipeline output"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
    theme: flatly
    highlight: tango
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath("../.."))

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)

# Create output directories if they don't exist
dir.create("analysis/output/corpus_descriptives/figures", showWarnings = FALSE, recursive = TRUE)
dir.create("analysis/output/corpus_descriptives/tables", showWarnings = FALSE, recursive = TRUE)
```

```{r libraries}
suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(knitr)
  library(kableExtra)
})
```

# Introduction

This report presents a **linguistically-informed** analysis of corpus statistics, organized around key phenomena relevant to null-subject research:

1. **Null Subject Analysis** — Where do missing subjects occur in English, and what explains them?
2. **Pronoun System** — The paradigmatic structure and functional distribution of overt pronouns
3. **Extraction and That-Trace** — Complementizer presence and filler-gap dependencies
4. **Supporting Analyses** — Verb finiteness, negation, and relative clauses

The analysis draws on 8 linguistic analyzers run across 6 text genres (BNC, CHILDES, Gutenberg, OpenSubtitles, SimpleWikipedia, Switchboard).

---

# Data Loading and Helper Functions

```{r data-setup}
# Genre colour palette (colorblind-friendly)
GENRE_COLORS <- c(
  BNC             = "#1f77b4",
  Gutenberg       = "#ff7f0e",
  OpenSubtitles   = "#2ca02c",
  SimpleWikipedia = "#d62728",
  Switchboard     = "#9467bd",
  CHILDES_child   = "#17becf",
  CHILDES_adult   = "#e377c2",
  CHILDES         = "#8c564b"
)

GENRE_ORDER <- c(
  "CHILDES_child", "CHILDES_adult", "CHILDES",
  "BNC", "Gutenberg", "OpenSubtitles", "SimpleWikipedia", "Switchboard"
)

# Spoken vs written classification
SPOKEN_GENRES <- c("CHILDES", "CHILDES_child", "CHILDES_adult", "OpenSubtitles", "Switchboard", "BNC")
WRITTEN_GENRES <- c("Gutenberg", "SimpleWikipedia")

ANALYZER_NAMES <- c(
  "pronoun_inventory", "expletives", "verb_finiteness", "clause_structure",
  "wh_questions", "relative_clauses", "negation", "that_trace"
)

SPLITS <- c("train_90M", "test_10M", "pull_10M")

load_analyzer_csv <- function(analyzer, split, base_dir = "analysis/output/corpus_descriptives/data") {
  path <- file.path(base_dir, split, paste0(analyzer, ".csv"))
  if (!file.exists(path)) return(NULL)
  read.csv(path, stringsAsFactors = FALSE)
}

load_all <- function() {
  all_data <- list()
  for (analyzer in ANALYZER_NAMES) {
    frames <- list()
    for (split in SPLITS) {
      df <- load_analyzer_csv(analyzer, split)
      if (!is.null(df)) frames[[split]] <- df
    }
    if (length(frames) > 0) all_data[[analyzer]] <- bind_rows(frames)
  }
  all_data
}

DATA <- load_all()

# Helper: filter to train_90M and specific genres
train_genre <- function(df) {
  df %>%
    filter(split == "train_90M", genre != "overall")
}

train_overall <- function(df) {
  df %>%
    filter(split == "train_90M", genre == "overall")
}

order_genres <- function(df) {
  present <- intersect(GENRE_ORDER, unique(df$genre))
  df %>% mutate(genre = factor(genre, levels = present))
}

# Check for CHILDES speaker split
SPEAKER_SPLIT_AVAILABLE <- any(DATA[["pronoun_inventory"]]$genre %in% c("CHILDES_child", "CHILDES_adult"))
```

```{r helper-functions}
#' Compute null-subject rate by grouping variables
#' Returns: overt, none, expletive counts + rates
compute_null_rate <- function(df, grouping_vars = NULL) {
  result <- df %>%
    filter(subject_status %in% c("overt", "none", "expletive"),
           !is.na(count))

  if (!is.null(grouping_vars) && length(grouping_vars) > 0) {
    result <- result %>%
      group_by(across(all_of(grouping_vars)), subject_status)
  } else {
    result <- result %>%
      group_by(subject_status)
  }

  result %>%
    summarise(n = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = subject_status, values_from = n, values_fill = 0) %>%
    mutate(
      # Ensure columns exist
      overt = if ("overt" %in% names(.)) overt else 0,
      none = if ("none" %in% names(.)) none else 0,
      expletive = if ("expletive" %in% names(.)) expletive else 0,
      total = overt + none + expletive,
      null_rate = none / total,
      expletive_rate = expletive / total,
      overt_rate = overt / total,
      # Adjusted null rate (excluding expletives from denominator)
      adj_null_rate = none / (overt + none)
    )
}

#' Format percentage for display
fmt_pct <- function(x, digits = 1) {
  sprintf(paste0("%.", digits, "f%%"), x * 100)
}

#' Format count with thousands separator
fmt_count <- function(x) {
  format(x, big.mark = ",", scientific = FALSE)
}

#' Standard theme for plots
theme_corpus <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "right",
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(color = "grey40")
    )
}

#' Save figure to output directory
save_figure <- function(plot, name, width = 10, height = 6) {
  ggsave(
    filename = paste0("analysis/output/corpus_descriptives/figures/", name, ".png"),
    plot = plot,
    width = width,
    height = height,
    dpi = 150
  )
}

#' Save table to output directory
save_table <- function(df, name) {
  write.csv(df,
            file = paste0("analysis/output/corpus_descriptives/tables/", name, ".csv"),
            row.names = FALSE)
}
```

**Data loaded:**

- **Analyzers:** `r paste(names(DATA), collapse = ", ")`
- **Splits:** `r paste(SPLITS[SPLITS %in% unique(bind_rows(DATA)$split)], collapse = ", ")`
- **CHILDES speaker split:** `r ifelse(SPEAKER_SPLIT_AVAILABLE, "Available (child/adult separated)", "Not yet available (awaiting pipeline re-run)")`

---

# Part 1: Executive Summary {.tabset}

Key findings across all phenomena from the 90M-token training corpus.

```{r executive-summary}
if (!is.null(DATA[["clause_structure"]])) {
  # Overall null-subject rate in finite ROOT clauses
  cs_overall <- train_overall(DATA[["clause_structure"]]) %>%
    filter(clause_dep == "ROOT")

  root_rates <- compute_null_rate(cs_overall)

  # By clause type
  cs_by_type <- train_overall(DATA[["clause_structure"]]) %>%
    filter(!is.na(clause_dep), clause_dep != "") %>%
    compute_null_rate(grouping_vars = "clause_dep")
}

if (!is.null(DATA[["expletives"]])) {
  exp_overall <- train_overall(DATA[["expletives"]]) %>%
    filter(!is.na(class), class != "")

  total_expletives <- sum(as.numeric(exp_overall$count), na.rm = TRUE)
}
```

## Key Metrics

```{r summary-metrics, results='asis'}
cat('<div class="row">\n')
cat('<div class="col-md-4">\n')
cat('<div class="panel panel-primary">\n')
cat('<div class="panel-heading"><strong>Null Subject Rate (ROOT)</strong></div>\n')
cat('<div class="panel-body" style="font-size: 24px; text-align: center;">\n')
if (exists("root_rates")) {
  cat(fmt_pct(root_rates$null_rate), '\n')
  cat('<br><small style="color: grey;">of all ROOT clauses</small>\n')
}
cat('</div></div></div>\n')

cat('<div class="col-md-4">\n')
cat('<div class="panel panel-info">\n')
cat('<div class="panel-heading"><strong>Expletive Rate (ROOT)</strong></div>\n')
cat('<div class="panel-body" style="font-size: 24px; text-align: center;">\n')
if (exists("root_rates")) {
  cat(fmt_pct(root_rates$expletive_rate), '\n')
  cat('<br><small style="color: grey;">it/there subjects</small>\n')
}
cat('</div></div></div>\n')

cat('<div class="col-md-4">\n')
cat('<div class="panel panel-success">\n')
cat('<div class="panel-heading"><strong>Overt Subject Rate (ROOT)</strong></div>\n')
cat('<div class="panel-body" style="font-size: 24px; text-align: center;">\n')
if (exists("root_rates")) {
  cat(fmt_pct(root_rates$overt_rate), '\n')
  cat('<br><small style="color: grey;">lexical or pronominal</small>\n')
}
cat('</div></div></div>\n')
cat('</div>\n')
```

## Subject Realization by Clause Type

```{r summary-clause-table}
if (exists("cs_by_type")) {
  clause_summary <- cs_by_type %>%
    arrange(desc(total)) %>%
    mutate(
      clause_type = case_when(
        clause_dep == "ROOT" ~ "ROOT (main clause)",
        clause_dep == "ccomp" ~ "ccomp (complement clause)",
        clause_dep == "advcl" ~ "advcl (adverbial clause)",
        clause_dep == "acl" ~ "acl (adjectival clause)",
        TRUE ~ clause_dep
      )
    ) %>%
    select(
      `Clause Type` = clause_type,
      Overt = overt,
      `Null (none)` = none,
      Expletive = expletive,
      Total = total,
      `Null Rate` = null_rate,
      `Adj. Null Rate` = adj_null_rate
    ) %>%
    mutate(
      `Null Rate` = fmt_pct(`Null Rate`),
      `Adj. Null Rate` = fmt_pct(`Adj. Null Rate`)
    )

  kable(clause_summary,
        caption = "Subject realization by clause type (train_90M, all genres)",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
    footnote(general = "Adj. Null Rate excludes expletives from denominator")

  save_table(clause_summary, "summary_clause_types")
}
```

## Interpretation

**What the null-subject rate tells us:**

English is a **non-pro-drop language** — overt subjects are required in most finite clauses. The ~6-7% null-subject rate in ROOT clauses represents:

1. **Imperatives** — "Go away!" (subject *you* is understood)
2. **Diary drop** — "Went to the store today" (colloquial subject omission)
3. **Fragments** — "Pretty good" (not full clauses)
4. **Parsing artifacts** — Some dependency misparses

**Why xcomp has high null rates:**

The `xcomp` (open clausal complement) relation covers infinitival and gerundive complements like "want to go" or "tried running." These contain **PRO** — the silent subject of non-finite clauses — which is grammatically required, not a gap.

---

# Part 2: Null Subject Analysis {.tabset}

The core analysis for understanding where English allows null subjects and why.

## 2.1 Overall Rates

```{r null-overall-rates}
if (!is.null(DATA[["clause_structure"]])) {
  # Compute rates for each clause type
  cs_train <- train_overall(DATA[["clause_structure"]]) %>%
    filter(!is.na(clause_dep), clause_dep != "",
           !is.na(subject_status), subject_status != "")

  null_by_clause <- cs_train %>%
    compute_null_rate(grouping_vars = "clause_dep") %>%
    arrange(desc(total))

  # Create visualization
  null_rates_long <- null_by_clause %>%
    select(clause_dep, overt_rate, null_rate, expletive_rate) %>%
    pivot_longer(
      cols = c(overt_rate, null_rate, expletive_rate),
      names_to = "status",
      values_to = "rate"
    ) %>%
    mutate(
      status = factor(status,
                      levels = c("overt_rate", "expletive_rate", "null_rate"),
                      labels = c("Overt", "Expletive", "Null"))
    )

  p_null_rates <- ggplot(null_rates_long,
                         aes(x = reorder(clause_dep, -rate), y = rate, fill = status)) +
    geom_col(position = "stack", width = 0.7) +
    scale_fill_manual(values = c("Overt" = "#2E86AB", "Expletive" = "#E8C547", "Null" = "#A23B72")) +
    scale_y_continuous(labels = percent_format()) +
    labs(
      title = "Subject Realization by Clause Type",
      subtitle = "Null subjects are rare in finite clauses (ROOT, ccomp, advcl) but common in non-finite (xcomp, acl)",
      x = "Clause Dependency Type",
      y = "Proportion of clauses",
      fill = "Subject Status"
    ) +
    theme_corpus()

  print(p_null_rates)
  save_figure(p_null_rates, "null_rates_by_clause")
}
```

### Rate Table

```{r null-rate-table}
if (exists("null_by_clause")) {
  null_table <- null_by_clause %>%
    mutate(
      finiteness = case_when(
        clause_dep %in% c("ROOT", "ccomp", "advcl") ~ "Typically finite",
        clause_dep %in% c("xcomp", "acl") ~ "Often non-finite",
        TRUE ~ "Mixed"
      )
    ) %>%
    arrange(finiteness, desc(total)) %>%
    select(
      `Clause Type` = clause_dep,
      `Finiteness` = finiteness,
      Overt = overt,
      Null = none,
      Expletive = expletive,
      Total = total,
      `Null %` = null_rate,
      `Overt %` = overt_rate
    ) %>%
    mutate(
      `Null %` = fmt_pct(`Null %`),
      `Overt %` = fmt_pct(`Overt %`)
    )

  kable(null_table,
        caption = "Subject realization rates by clause type",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
    row_spec(which(null_table$Finiteness == "Often non-finite"), background = "#f8f9fa")
}
```

## 2.2 By Genre

```{r null-by-genre}
if (!is.null(DATA[["clause_structure"]])) {
  # Compute null rates by genre for ROOT clauses only
  cs_genre <- train_genre(DATA[["clause_structure"]]) %>%
    filter(clause_dep == "ROOT",
           !is.na(subject_status), subject_status != "") %>%
    order_genres()

  null_by_genre <- cs_genre %>%
    compute_null_rate(grouping_vars = "genre") %>%
    mutate(
      modality = ifelse(genre %in% SPOKEN_GENRES, "Spoken/Dialogue", "Written"),
      genre = factor(genre, levels = GENRE_ORDER[GENRE_ORDER %in% genre])
    )

  # Visualization: null rate by genre with modality grouping
  p_genre_null <- ggplot(null_by_genre,
                         aes(x = genre, y = null_rate, fill = modality)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = fmt_pct(null_rate)), vjust = -0.5, size = 3) +
    scale_fill_manual(values = c("Spoken/Dialogue" = "#2E86AB", "Written" = "#A23B72")) +
    scale_y_continuous(labels = percent_format(), limits = c(0, max(null_by_genre$null_rate) * 1.15)) +
    labs(
      title = "Null Subject Rate in ROOT Clauses by Genre",
      subtitle = "Spoken/dialogue genres show higher null-subject rates (diary drop, fragments)",
      x = NULL,
      y = "Null subject rate",
      fill = "Modality"
    ) +
    theme_corpus()

  print(p_genre_null)
  save_figure(p_genre_null, "null_rate_by_genre")
}
```

### Genre Comparison Table

```{r genre-table}
if (exists("null_by_genre")) {
  genre_table <- null_by_genre %>%
    arrange(desc(null_rate)) %>%
    select(
      Genre = genre,
      Modality = modality,
      Overt = overt,
      Null = none,
      Expletive = expletive,
      Total = total,
      `Null Rate` = null_rate,
      `Expletive Rate` = expletive_rate
    ) %>%
    mutate(
      `Null Rate` = fmt_pct(`Null Rate`),
      `Expletive Rate` = fmt_pct(`Expletive Rate`)
    )

  kable(genre_table,
        caption = "ROOT clause subject realization by genre",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(genre_table, "null_rate_by_genre")
}
```

### Interpretation

**Genre patterns align with linguistic expectations:**

- **CHILDES** has the highest null rate (~10-12%): Child-directed speech includes many imperatives, sentence fragments, and informal register
- **OpenSubtitles** (movie dialogue) shows elevated rates: Colloquial speech, truncated responses
- **SimpleWikipedia** has the lowest: Formal expository prose requires explicit subjects

## 2.3 Control/Raising Verb Inventory

Non-finite complements (xcomp) are licensed by **control verbs** (the subject is interpreted) and **raising verbs** (the subject raises from the embedded clause).

```{r control-raising-verbs}
if (!is.null(DATA[["clause_structure"]])) {
  # Extract verb lemmas from clause_structure (these are the matrix verbs)
  verb_data <- train_overall(DATA[["clause_structure"]]) %>%
    filter(!is.na(verb_lemma), verb_lemma != "") %>%
    group_by(verb_lemma) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(count))

  # Classify verbs (based on linguistic knowledge)
  control_verbs <- c("want", "try", "decide", "hope", "plan", "refuse", "promise",
                     "agree", "offer", "threaten", "choose", "wish", "intend",
                     "manage", "fail", "pretend", "learn", "prepare")
  raising_verbs <- c("seem", "appear", "happen", "tend", "turn", "prove")
  modal_like <- c("have", "be", "got", "need", "ought", "able", "going", "hafta",
                  "wanna", "gonna", "gotta", "used", "about", "likely")

  verb_data <- verb_data %>%
    mutate(
      verb_class = case_when(
        verb_lemma %in% control_verbs ~ "Control",
        verb_lemma %in% raising_verbs ~ "Raising",
        verb_lemma %in% modal_like ~ "Modal-like",
        TRUE ~ "Other"
      )
    )

  # Top 30 verbs
  top_verbs <- verb_data %>%
    head(30) %>%
    mutate(
      rank = row_number(),
      verb_display = paste0(rank, ". ", verb_lemma)
    )

  p_verbs <- ggplot(top_verbs,
                    aes(x = reorder(verb_lemma, count), y = count, fill = verb_class)) +
    geom_col(width = 0.7) +
    coord_flip() +
    scale_fill_manual(values = c("Control" = "#2E86AB", "Raising" = "#E8C547",
                                  "Modal-like" = "#8c564b", "Other" = "#999999")) +
    scale_y_continuous(labels = label_comma()) +
    labs(
      title = "Top 30 Matrix Verbs Licensing Infinitival Complements",
      subtitle = "Control verbs (want, try) and modal-like verbs dominate; raising verbs (seem, appear) are rarer",
      x = NULL,
      y = "Count",
      fill = "Verb Class"
    ) +
    theme_corpus() +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

  print(p_verbs)
  save_figure(p_verbs, "control_raising_verbs")
}
```

### Verb Class Summary

```{r verb-class-summary}
if (exists("verb_data")) {
  class_summary <- verb_data %>%
    filter(verb_class != "Other") %>%
    group_by(verb_class) %>%
    summarise(
      `Verb Types` = n(),
      `Total Tokens` = sum(count),
      `Top Examples` = paste(head(verb_lemma, 5), collapse = ", "),
      .groups = "drop"
    ) %>%
    arrange(desc(`Total Tokens`))

  kable(class_summary,
        caption = "Infinitival complement verbs by class",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(class_summary, "verb_class_summary")
}
```

### Interpretation

**Control vs Raising:**

- **Control verbs** (want, try, decide, hope, promise...) — The matrix subject is understood as the embedded subject. "John wants to leave" = John is the leaver.
- **Raising verbs** (seem, appear, happen, tend...) — The subject originates in the embedded clause and raises. "John seems to be happy" ← "to be happy" originally had John as its subject.
- **Modal-like verbs** (have to, be going to, need to...) — Grammaticalized auxiliaries behaving like modals.

Raising verbs are much rarer than control verbs in natural language use.

## 2.4 Expletive vs True Null

Expletive subjects (*it*, *there*) are not referential — they fill the subject position without denoting an entity.

```{r expletive-analysis}
if (!is.null(DATA[["expletives"]])) {
  exp_data <- train_overall(DATA[["expletives"]]) %>%
    filter(!is.na(class), class != "")

  # Class counts
  exp_by_class <- exp_data %>%
    group_by(class) %>%
    summarise(
      count = sum(as.numeric(count), na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(desc(count)) %>%
    mutate(
      proportion = count / sum(count),
      description = case_when(
        class == "existential" ~ "There is/are... (existential)",
        class == "weather" ~ "It rains, snows... (weather)",
        class == "raising" ~ "It seems that... (raising)",
        TRUE ~ class
      )
    )

  p_exp <- ggplot(exp_by_class, aes(x = reorder(class, count), y = count, fill = class)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = paste0(fmt_count(count), " (", fmt_pct(proportion), ")")),
              hjust = -0.1, size = 3.5) +
    coord_flip() +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(labels = label_comma(), limits = c(0, max(exp_by_class$count) * 1.3)) +
    labs(
      title = "Expletive Subject Types",
      subtitle = "Existential 'there' dominates; weather 'it' and raising 'it' are rare",
      x = NULL,
      y = "Count"
    ) +
    theme_corpus() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 0))

  print(p_exp)
  save_figure(p_exp, "expletive_types")
}
```

### Top Expletive Verbs

```{r expletive-verbs}
if (!is.null(DATA[["expletives"]])) {
  exp_verbs <- train_overall(DATA[["expletives"]]) %>%
    filter(!is.na(class), class != "", !is.na(verb_lemma), verb_lemma != "") %>%
    group_by(class, verb_lemma) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    arrange(class, desc(count)) %>%
    group_by(class) %>%
    slice_head(n = 10) %>%
    ungroup()

  exp_verbs_wide <- exp_verbs %>%
    group_by(class) %>%
    mutate(rank = row_number()) %>%
    pivot_wider(
      id_cols = rank,
      names_from = class,
      values_from = c(verb_lemma, count)
    )

  # Just show existential (by far the most common)
  exp_existential <- exp_verbs %>%
    filter(class == "existential") %>%
    select(Verb = verb_lemma, Count = count) %>%
    mutate(Example = case_when(
      Verb == "be" ~ "There is/are/was...",
      Verb == "go" ~ "There goes the...",
      Verb == "come" ~ "There came a...",
      Verb == "seem" ~ "There seems to be...",
      Verb == "exist" ~ "There exists...",
      TRUE ~ ""
    ))

  kable(exp_existential,
        caption = "Top 10 verbs in existential constructions",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
}
```

### Adjusted Null Rate (Excluding Expletives)

```{r adjusted-null-rate}
if (!is.null(DATA[["clause_structure"]])) {
  # Compare raw vs adjusted null rates by genre
  adj_comparison <- train_genre(DATA[["clause_structure"]]) %>%
    filter(clause_dep == "ROOT", !is.na(subject_status)) %>%
    order_genres() %>%
    compute_null_rate(grouping_vars = "genre") %>%
    select(genre, null_rate, adj_null_rate, expletive_rate) %>%
    pivot_longer(cols = c(null_rate, adj_null_rate),
                 names_to = "measure", values_to = "rate") %>%
    mutate(
      measure = factor(measure,
                       levels = c("null_rate", "adj_null_rate"),
                       labels = c("Raw Null Rate", "Adjusted (excl. expletives)"))
    )

  p_adj <- ggplot(adj_comparison, aes(x = genre, y = rate, fill = measure)) +
    geom_col(position = "dodge", width = 0.7) +
    scale_fill_manual(values = c("Raw Null Rate" = "#A23B72", "Adjusted (excl. expletives)" = "#2E86AB")) +
    scale_y_continuous(labels = percent_format()) +
    labs(
      title = "Raw vs Adjusted Null Subject Rate by Genre",
      subtitle = "Adjusted rate = none / (overt + none), excluding expletives from denominator",
      x = NULL,
      y = "Rate",
      fill = NULL
    ) +
    theme_corpus()

  print(p_adj)
  save_figure(p_adj, "adjusted_null_rate")
}
```

---

# Part 3: Pronoun System Characterization {.tabset}

## 3.1 Paradigmatic Inventory

The English personal pronoun paradigm organized by person and number.

```{r pronoun-paradigm}
if (!is.null(DATA[["pronoun_inventory"]])) {
  fn_col <- if ("function." %in% names(DATA[["pronoun_inventory"]])) "function." else "function"

  pi_overall <- train_overall(DATA[["pronoun_inventory"]])

  # Calculate totals for proportions
  total_pronouns <- sum(as.numeric(pi_overall$count), na.rm = TRUE)

  # Personal pronouns with person values
  pi_paradigm <- pi_overall %>%
    filter(!is.na(person), person != "", person %in% c(1, 2, 3)) %>%
    mutate(
      person = as.integer(person),
      person_label = case_when(
        person == 1 ~ "1st",
        person == 2 ~ "2nd",
        person == 3 ~ "3rd"
      ),
      number_label = case_when(
        person == 2 ~ "Sg/Pl",  # 2nd person is syncretic
        number == "Sing" ~ "Singular",
        number == "Plur" ~ "Plural",
        TRUE ~ "Other"
      )
    )

  # Aggregate by person × number
  paradigm_counts <- pi_paradigm %>%
    group_by(person_label, number_label) %>%
    summarise(
      count = sum(as.numeric(count), na.rm = TRUE),
      top_lemmas = paste(head(lemma[order(-as.numeric(count))], 3), collapse = ", "),
      .groups = "drop"
    ) %>%
    mutate(
      pct = count / total_pronouns * 100
    )

  # Create wide format table
  paradigm_wide <- paradigm_counts %>%
    select(person_label, number_label, count, pct, top_lemmas) %>%
    mutate(
      cell = paste0(fmt_count(count), "\n(", sprintf("%.1f%%", pct), ")\n", top_lemmas)
    )
}
```

```{r pronoun-paradigm-table, results='asis'}
if (exists("paradigm_counts")) {
  # Create nice HTML table
  cat('<table class="table table-striped table-hover" style="width: auto !important; margin-left: auto; margin-right: auto;">
<caption><strong>Personal Pronoun Paradigm (Subject Position)</strong></caption>
<thead>
<tr><th style="text-align:left;">Person</th><th style="text-align:center; width:200px;">Singular</th><th style="text-align:center; width:200px;">Plural</th></tr>
</thead>
<tbody>\n')

  for (p in c("1st", "2nd", "3rd")) {
    cat('<tr><td style="text-align:left; vertical-align:top;"><strong>', p, '</strong></td>')

    if (p == "2nd") {
      # Syncretic 2nd person
      row <- paradigm_counts %>% filter(person_label == "2nd")
      if (nrow(row) > 0) {
        cat('<td colspan="2" style="text-align:center; vertical-align:top;"><em>you</em> (syncretic)<br>',
            fmt_count(row$count[1]), '<br>',
            '<small>', sprintf("%.1f%%", row$pct[1]), ' of pronouns</small></td>')
      }
    } else {
      # Singular
      sg <- paradigm_counts %>% filter(person_label == p, number_label == "Singular")
      if (nrow(sg) > 0) {
        cat('<td style="text-align:center; vertical-align:top;">',
            fmt_count(sg$count), '<br>',
            '<small>', sprintf("%.1f%%", sg$pct), '</small><br>',
            '<small style="color:grey;">', sg$top_lemmas, '</small></td>')
      } else {
        cat('<td style="text-align:center;">—</td>')
      }

      # Plural
      pl <- paradigm_counts %>% filter(person_label == p, number_label == "Plural")
      if (nrow(pl) > 0) {
        cat('<td style="text-align:center; vertical-align:top;">',
            fmt_count(pl$count), '<br>',
            '<small>', sprintf("%.1f%%", pl$pct), '</small><br>',
            '<small style="color:grey;">', pl$top_lemmas, '</small></td>')
      } else {
        cat('<td style="text-align:center;">—</td>')
      }
    }
    cat('</tr>\n')
  }

  cat('</tbody></table>\n')
  cat('<p><small><em>Each cell shows: count, % of all pronouns, top lemmas</em></small></p>\n')
}
```

## 3.2 Functional Distribution

```{r pronoun-function}
if (!is.null(DATA[["pronoun_inventory"]])) {
  fn_col <- if ("function." %in% names(pi_overall)) "function." else "function"

  fn_summary <- pi_overall %>%
    rename(fn = all_of(fn_col)) %>%
    filter(fn %in% c("subject", "object")) %>%
    group_by(fn) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    mutate(pct = count / sum(count) * 100)

  kable(fn_summary %>%
          mutate(count = fmt_count(count), pct = sprintf("%.1f%%", pct)) %>%
          rename(Function = fn, Count = count, Percentage = pct),
        caption = "Pronoun distribution by grammatical function") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
}
```

### Person Distribution

```{r pronoun-person-dist}
if (!is.null(DATA[["pronoun_inventory"]])) {
  person_dist <- pi_overall %>%
    filter(!is.na(person), person %in% c(1, 2, 3)) %>%
    mutate(person = as.integer(person)) %>%
    group_by(person) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    mutate(
      person_label = paste0(person, ifelse(person == 1, "st", ifelse(person == 2, "nd", "rd")), " person"),
      pct = count / sum(count) * 100
    )

  p_person <- ggplot(person_dist, aes(x = factor(person), y = count, fill = factor(person))) +
    geom_col(width = 0.6) +
    geom_text(aes(label = sprintf("%s\n(%.1f%%)", fmt_count(count), pct)), vjust = -0.3, size = 3.5) +
    scale_fill_manual(values = c("1" = "#2E86AB", "2" = "#E8C547", "3" = "#A23B72")) +
    scale_y_continuous(labels = label_comma(), limits = c(0, max(person_dist$count) * 1.15)) +
    labs(
      title = "Pronoun Distribution by Person",
      subtitle = "1st and 2nd person dominate — consistent with dialogue-heavy corpora",
      x = "Person",
      y = "Count"
    ) +
    theme_corpus() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 0))

  print(p_person)
  save_figure(p_person, "pronoun_person_distribution")
}
```

## 3.3 Genre Variation

```{r pronoun-by-genre}
if (!is.null(DATA[["pronoun_inventory"]])) {
  pi_genre <- train_genre(DATA[["pronoun_inventory"]]) %>%
    order_genres() %>%
    filter(!is.na(person), person %in% c(1, 2, 3)) %>%
    mutate(person = as.integer(person)) %>%
    group_by(genre, person) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    group_by(genre) %>%
    mutate(pct = count / sum(count) * 100) %>%
    ungroup()

  p_genre_person <- ggplot(pi_genre, aes(x = genre, y = pct, fill = factor(person))) +
    geom_col(position = "stack", width = 0.7) +
    scale_fill_manual(values = c("1" = "#2E86AB", "2" = "#E8C547", "3" = "#A23B72"),
                      labels = c("1st", "2nd", "3rd")) +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    labs(
      title = "Person Distribution by Genre",
      subtitle = "CHILDES and dialogue corpora have more 1st/2nd person; Wikipedia has more 3rd person",
      x = NULL,
      y = "Percentage of pronouns",
      fill = "Person"
    ) +
    theme_corpus()

  print(p_genre_person)
  save_figure(p_genre_person, "pronoun_person_by_genre")
}
```

### Interpretation

**Genre-driven variation:**

- **CHILDES** (child-directed speech): Heavy use of 1st/2nd person — "I", "you" — in interaction
- **OpenSubtitles** (dialogue): Similar pattern to CHILDES
- **SimpleWikipedia** (encyclopedia): 3rd person dominates — "he", "she", "it", "they" — for reference
- **Gutenberg** (fiction): Mix of dialogue (1st/2nd) and narration (3rd)

This variation is expected and validates the genre diversity of the corpus.

---

# Part 4: Extraction and That-Trace {.tabset}

Analysis of complement clauses, complementizer presence, and extraction (that-trace effects).

## 4.1 Bridge Verb Inventory

Bridge verbs (think, say, know, believe...) take clausal complements and allow extraction.

```{r bridge-verbs}
if (!is.null(DATA[["that_trace"]])) {
  tt_data <- train_overall(DATA[["that_trace"]])

  bridge_verbs <- tt_data %>%
    filter(!is.na(verb_lemma), verb_lemma != "") %>%
    group_by(verb_lemma) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(count)) %>%
    head(20)

  p_bridge <- ggplot(bridge_verbs,
                     aes(x = reorder(verb_lemma, count), y = count)) +
    geom_col(fill = "#2E86AB", width = 0.7) +
    coord_flip() +
    scale_y_continuous(labels = label_comma()) +
    labs(
      title = "Top 20 Bridge (Complement-Taking) Verbs",
      subtitle = "Verbs that embed 'that'-clauses: 'think', 'say', 'know' dominate",
      x = NULL,
      y = "Count"
    ) +
    theme_corpus() +
    theme(axis.text.x = element_text(angle = 0))

  print(p_bridge)
  save_figure(p_bridge, "bridge_verbs")
}
```

## 4.2 Complementizer Presence

The complementizer "that" can be omitted in English ("I think (that) he left").

```{r complementizer-rate}
if (!is.null(DATA[["that_trace"]])) {
  # Get complementizer rate data
  comp_data <- tt_data %>%
    filter(metric == "complementizer_rate") %>%
    mutate(
      with_that = as.numeric(with_that),
      without_that = as.numeric(without_that),
      total = with_that + without_that,
      that_rate = with_that / total
    )

  # By genre
  comp_genre <- train_genre(DATA[["that_trace"]]) %>%
    filter(metric == "complementizer_rate") %>%
    mutate(
      with_that = as.numeric(with_that),
      without_that = as.numeric(without_that),
      total = with_that + without_that,
      that_rate = with_that / total
    ) %>%
    order_genres()

  if (nrow(comp_genre) > 0) {
    p_comp <- ggplot(comp_genre, aes(x = genre, y = that_rate, fill = genre)) +
      geom_col(width = 0.7) +
      geom_text(aes(label = fmt_pct(that_rate)), vjust = -0.5, size = 3) +
      scale_fill_manual(values = GENRE_COLORS) +
      scale_y_continuous(labels = percent_format(), limits = c(0, max(comp_genre$that_rate) * 1.15)) +
      labs(
        title = "Complementizer 'that' Retention Rate by Genre",
        subtitle = "Higher rates in formal written genres; lower in spoken/colloquial",
        x = NULL,
        y = "'that' retention rate",
        fill = NULL
      ) +
      theme_corpus() +
      theme(legend.position = "none")

    print(p_comp)
    save_figure(p_comp, "complementizer_rate_by_genre")
  }
}
```

### Interpretation

**Complementizer deletion patterns:**

- **Formal written prose** (SimpleWikipedia): Higher *that* retention — "The study shows **that** results were..."
- **Spoken/colloquial** (CHILDES, OpenSubtitles): More *that* deletion — "I think Ø he's right"
- This reflects register differences: formal = more explicit marking; informal = more ellipsis

## 4.3 That-Trace Effects

The **that-trace effect** is a classic constraint in English syntax: subject extraction from an embedded clause is degraded when the complementizer "that" is present.

| Condition | Example | Judgment |
|-----------|---------|----------|
| Subject extraction, no *that* | "Who do you think __ left?" | ✓ Grammatical |
| Subject extraction, with *that* | "Who do you think *that* __ left?" | * Ungrammatical |
| Object extraction, no *that* | "What do you think he ate __?" | ✓ Grammatical |
| Object extraction, with *that* | "What do you think *that* he ate __?" | ✓ Grammatical |

The asymmetry is striking: *that* is fine with object extraction but blocks subject extraction. This is attributed to the **Empty Category Principle** (ECP) in Government & Binding theory, or **anti-locality** constraints in Minimalism.

### Contingency Table

```{r that-trace-contingency}
if (!is.null(DATA[["that_trace"]])) {
  # Build the 2×2 contingency table for extraction × complementizer
  tt_extraction <- tt_data %>%
    filter(!is.na(comp_present), comp_present != "",
           !is.na(extraction_type), extraction_type %in% c("subject", "object")) %>%
    group_by(comp_present, extraction_type) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop")

  if (nrow(tt_extraction) > 0) {
    # Create labeled contingency table
    tt_table <- tt_extraction %>%
      mutate(
        comp_label = ifelse(comp_present == "True", "With THAT", "Without THAT"),
        extraction_label = ifelse(extraction_type == "subject", "Subject Extraction", "Object Extraction")
      ) %>%
      select(comp_label, extraction_label, count) %>%
      pivot_wider(names_from = extraction_label, values_from = count, values_fill = 0) %>%
      mutate(Total = `Subject Extraction` + `Object Extraction`)

    # Add row totals
    tt_table_display <- tt_table %>%
      bind_rows(
        tibble(
          comp_label = "Total",
          `Subject Extraction` = sum(tt_table$`Subject Extraction`),
          `Object Extraction` = sum(tt_table$`Object Extraction`),
          Total = sum(tt_table$Total)
        )
      )

    kable(tt_table_display,
          col.names = c("Complementizer", "Subject Extraction", "Object Extraction", "Total"),
          caption = "Extraction Type × Complementizer Presence",
          format.args = list(big.mark = ",")) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
      row_spec(3, bold = TRUE)
  }
}
```

### Statistical Analysis

```{r that-trace-stats, results='asis'}
if (exists("tt_extraction") && nrow(tt_extraction) > 0) {
  # Extract cell values
  subj_that <- tt_extraction %>% filter(comp_present == "True", extraction_type == "subject") %>% pull(count)
  subj_no_that <- tt_extraction %>% filter(comp_present == "False", extraction_type == "subject") %>% pull(count)
  obj_that <- tt_extraction %>% filter(comp_present == "True", extraction_type == "object") %>% pull(count)
  obj_no_that <- tt_extraction %>% filter(comp_present == "False", extraction_type == "object") %>% pull(count)

  # Handle missing values
  if (length(subj_that) == 0) subj_that <- 0
  if (length(subj_no_that) == 0) subj_no_that <- 0
  if (length(obj_that) == 0) obj_that <- 0
  if (length(obj_no_that) == 0) obj_no_that <- 0

  # Compute rates
  subj_total <- subj_that + subj_no_that
  obj_total <- obj_that + obj_no_that
  that_total <- subj_that + obj_that
  no_that_total <- subj_no_that + obj_no_that

  # Rate of "that" usage by extraction type
  that_rate_subj <- subj_that / subj_total
  that_rate_obj <- obj_that / obj_total

  # Subject rate by complementizer condition
  subj_rate_that <- subj_that / that_total
  subj_rate_no_that <- subj_no_that / no_that_total

  # Odds ratio: (subj|no-that / obj|no-that) / (subj|that / obj|that)
  # OR > 1 means subject extraction is disfavored when "that" is present
  if (subj_that > 0 && obj_that > 0) {
    odds_ratio <- (subj_no_that * obj_that) / (subj_that * obj_no_that)
    log_or <- log(odds_ratio)
    se_log_or <- sqrt(1/subj_that + 1/subj_no_that + 1/obj_that + 1/obj_no_that)
    ci_lower <- exp(log_or - 1.96 * se_log_or)
    ci_upper <- exp(log_or + 1.96 * se_log_or)
  }

  cat('<div class="panel panel-info">\n')
  cat('<div class="panel-heading"><strong>That-Trace Effect: Key Statistics</strong></div>\n')
  cat('<div class="panel-body">\n')
  cat('<table class="table table-condensed" style="width: auto;">\n')

  cat('<tr><td><strong>Complementizer rate in subject extraction:</strong></td>')
  cat(sprintf('<td>%s (%d / %d)</td></tr>\n', fmt_pct(that_rate_subj), subj_that, subj_total))

  cat('<tr><td><strong>Complementizer rate in object extraction:</strong></td>')
  cat(sprintf('<td>%s (%d / %d)</td></tr>\n', fmt_pct(that_rate_obj), obj_that, obj_total))

  cat('<tr><td><strong>Subject extraction rate with "that":</strong></td>')
  cat(sprintf('<td>%s (%d / %d)</td></tr>\n', fmt_pct(subj_rate_that), subj_that, that_total))

  cat('<tr><td><strong>Subject extraction rate without "that":</strong></td>')
  cat(sprintf('<td>%s (%d / %d)</td></tr>\n', fmt_pct(subj_rate_no_that), subj_no_that, no_that_total))

  if (exists("odds_ratio")) {
    cat('<tr><td><strong>Odds ratio (that-trace effect):</strong></td>')
    cat(sprintf('<td>%.2f [95%% CI: %.2f – %.2f]</td></tr>\n', odds_ratio, ci_lower, ci_upper))
  }

  cat('</table>\n')
  cat('</div></div>\n')
}
```

### Visualization: The That-Trace Asymmetry

```{r that-trace-viz-improved}
if (exists("tt_extraction") && nrow(tt_extraction) > 0) {
  # Compute rates within each complementizer condition
  tt_rates <- tt_extraction %>%
    mutate(
      comp_label = factor(
        ifelse(comp_present == "True", "With 'that'", "Without 'that'"),
        levels = c("Without 'that'", "With 'that'")
      )
    ) %>%
    group_by(comp_label) %>%
    mutate(
      total = sum(count),
      rate = count / total
    ) %>%
    ungroup() %>%
    filter(extraction_type == "subject")  # Focus on subject extraction rate

  p_asymmetry <- ggplot(tt_rates, aes(x = comp_label, y = rate, fill = comp_label)) +
    geom_col(width = 0.5) +
    geom_text(aes(label = sprintf("%s\n(n=%s)", fmt_pct(rate), fmt_count(count))),
              vjust = -0.3, size = 3.5) +
    scale_fill_manual(values = c("Without 'that'" = "#2E86AB", "With 'that'" = "#A23B72")) +
    scale_y_continuous(labels = percent_format(), limits = c(0, max(tt_rates$rate) * 1.3)) +
    labs(
      title = "Subject Extraction Rate by Complementizer Condition",
      subtitle = sprintf("Subject extraction drops from %.1f%% to %.1f%% when 'that' is present — the that-trace effect",
                        tt_rates$rate[tt_rates$comp_label == "Without 'that'"] * 100,
                        tt_rates$rate[tt_rates$comp_label == "With 'that'"] * 100),
      x = NULL,
      y = "Subject Extraction Rate\n(among all extractions)",
      fill = NULL
    ) +
    theme_corpus() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 0, size = 12))

  print(p_asymmetry)
  save_figure(p_asymmetry, "that_trace_asymmetry")
}
```

### That-Trace "Violations" in the Corpus

```{r that-trace-violations, results='asis'}
if (exists("subj_that") && subj_that > 0) {
  cat(sprintf('\n<div class="alert alert-warning">\n'))
  cat(sprintf('<strong>Potential That-Trace Violations: %s cases</strong>\n', fmt_count(subj_that)))
  cat('\n\nThese are instances where the parser detected subject extraction co-occurring with an overt complementizer "that".\n')
  cat('</div>\n\n')

  cat('**What might these cases represent?**\n\n')
  cat('1. **Parsing errors** — The dependency parser may misidentify extraction type or complementizer status\n')
  cat('2. **Resumptive pronouns** — "Who do you think that *they* left?" (not true extraction)\n')
  cat('3. **Dialectal variation** — Some English dialects tolerate that-trace more than others\n')
  cat('4. **Heavy NP effects** — Long or complex extracted phrases may license "that"\n')
  cat('5. **Genre effects** — Spoken language may be more permissive\n\n')
}
```

### That-Trace Violations by Genre

```{r that-trace-by-genre}
if (!is.null(DATA[["that_trace"]])) {
  # Compute violations by genre
  tt_by_genre <- train_genre(DATA[["that_trace"]]) %>%
    filter(!is.na(comp_present), comp_present != "",
           !is.na(extraction_type), extraction_type %in% c("subject", "object")) %>%
    order_genres() %>%
    group_by(genre, comp_present, extraction_type) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop")

  # Extract violations (subject + that) by genre
  violations_by_genre <- tt_by_genre %>%
    filter(comp_present == "True", extraction_type == "subject") %>%
    select(genre, violations = count)

  # Get total subject extractions by genre for rate calculation
  subj_total_by_genre <- tt_by_genre %>%
    filter(extraction_type == "subject") %>%
    group_by(genre) %>%
    summarise(total_subj = sum(count), .groups = "drop")

  # Compute violation rate
  violation_rates <- violations_by_genre %>%
    left_join(subj_total_by_genre, by = "genre") %>%
    mutate(
      violation_rate = violations / total_subj,
      genre = factor(genre, levels = GENRE_ORDER[GENRE_ORDER %in% genre])
    ) %>%
    arrange(genre)

  if (nrow(violation_rates) > 0) {
    p_violations <- ggplot(violation_rates,
                           aes(x = genre, y = violations, fill = genre)) +
      geom_col(width = 0.7) +
      geom_text(aes(label = sprintf("%d\n(%.1f%%)", violations, violation_rate * 100)),
                vjust = -0.3, size = 3) +
      scale_fill_manual(values = GENRE_COLORS) +
      scale_y_continuous(limits = c(0, max(violation_rates$violations) * 1.3)) +
      labs(
        title = "That-Trace 'Violations' by Genre",
        subtitle = "Count (and rate among all subject extractions)",
        x = NULL,
        y = "Number of Violations",
        fill = NULL
      ) +
      theme_corpus() +
      theme(legend.position = "none")

    print(p_violations)
    save_figure(p_violations, "that_trace_violations_by_genre")
  }
}
```

### Violation Rate Table

```{r violation-table}
if (exists("violation_rates") && nrow(violation_rates) > 0) {
  violation_display <- violation_rates %>%
    arrange(desc(violation_rate)) %>%
    mutate(
      violation_rate = fmt_pct(violation_rate)
    ) %>%
    select(
      Genre = genre,
      `Violations (that + subj)` = violations,
      `Total Subject Extractions` = total_subj,
      `Violation Rate` = violation_rate
    )

  kable(violation_display,
        caption = "That-trace violations by genre",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(violation_display, "that_trace_violations")
}
```

### Interpretation

```{r that-trace-interpretation, results='asis'}
if (exists("odds_ratio")) {
  cat('**The that-trace effect is robustly attested in this corpus:**\n\n')

  if (odds_ratio > 5) {
    cat(sprintf('- The odds ratio of **%.1f** indicates a very strong effect — subject extraction is dramatically less likely when "that" is present.\n', odds_ratio))
  } else if (odds_ratio > 2) {
    cat(sprintf('- The odds ratio of **%.1f** indicates a moderate effect.\n', odds_ratio))
  } else {
    cat(sprintf('- The odds ratio of **%.1f** indicates a weak effect, possibly due to parsing noise.\n', odds_ratio))
  }

  cat(sprintf('- Only **%.1f%%** of subject extractions occur with "that", compared to **%.1f%%** of object extractions.\n',
              that_rate_subj * 100, that_rate_obj * 100))
  cat(sprintf('- The **%s violations** represent cases to investigate — likely a mix of parsing errors and genuine variation.\n\n',
              fmt_count(subj_that)))

  cat('**Linguistic significance:**\n\n')
  cat('The that-trace effect demonstrates that extraction in English is sensitive to local syntactic context. ')
  cat('The constraint is categorical in standard judgments but shows some "leakage" in naturalistic data, ')
  cat('which may reflect processing pressures, dialectal variation, or parser limitations.\n')
}
```

---

# Part 5: Supporting Analyses {.tabset}

## 5.1 Verb Finiteness

Distribution of finite vs non-finite verb forms across clause positions.

```{r verb-finiteness}
if (!is.null(DATA[["verb_finiteness"]])) {
  vf_data <- train_overall(DATA[["verb_finiteness"]]) %>%
    filter(!is.na(verb_form), verb_form != "",
           !is.na(clause_position), clause_position != "")

  vf_summary <- vf_data %>%
    group_by(verb_form, clause_position) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    mutate(
      finiteness = case_when(
        verb_form == "Fin" ~ "Finite",
        verb_form %in% c("Inf", "Part", "Ger") ~ "Non-finite",
        TRUE ~ "Other"
      )
    )

  p_vf <- ggplot(vf_summary,
                 aes(x = clause_position, y = count, fill = verb_form)) +
    geom_col(position = "stack", width = 0.7) +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(labels = label_comma()) +
    labs(
      title = "Verb Form Distribution by Clause Position",
      subtitle = "ROOT clauses are primarily finite; xcomp is primarily infinitival",
      x = "Clause Position",
      y = "Count",
      fill = "Verb Form"
    ) +
    theme_corpus()

  print(p_vf)
  save_figure(p_vf, "verb_finiteness_distribution")
}
```

### Finiteness by Clause Type Table

```{r vf-table}
if (exists("vf_summary")) {
  vf_wide <- vf_summary %>%
    pivot_wider(
      id_cols = clause_position,
      names_from = verb_form,
      values_from = count,
      values_fill = 0
    ) %>%
    mutate(Total = Fin + Inf + Part + Ger) %>%
    arrange(desc(Total))

  kable(vf_wide,
        caption = "Verb forms by clause position",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(vf_wide, "verb_finiteness_by_position")
}
```

## 5.2 Negation Patterns

Distribution of negation position (pre-verbal vs post-verbal) and subject realization.

```{r negation-analysis}
if (!is.null(DATA[["negation"]])) {
  neg_data <- train_overall(DATA[["negation"]]) %>%
    filter(!is.na(position), position != "",
           !is.na(subject_status), subject_status != "")

  neg_summary <- neg_data %>%
    group_by(position, subject_status) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop")

  p_neg <- ggplot(neg_summary,
                  aes(x = position, y = count, fill = subject_status)) +
    geom_col(position = "dodge", width = 0.7) +
    scale_fill_manual(values = c("overt" = "#2E86AB", "expletive" = "#E8C547",
                                  "none" = "#A23B72", "n/a" = "#999999")) +
    scale_y_continuous(labels = label_comma()) +
    labs(
      title = "Negation Position × Subject Realization",
      subtitle = "Pre-verbal negation (don't, won't) is most common; null subjects rare in negated clauses",
      x = "Negation Position",
      y = "Count",
      fill = "Subject Status"
    ) +
    theme_corpus() +
    theme(axis.text.x = element_text(angle = 0))

  print(p_neg)
  save_figure(p_neg, "negation_patterns")
}
```

### Negation Table

```{r negation-table}
if (exists("neg_summary")) {
  neg_wide <- neg_summary %>%
    pivot_wider(names_from = subject_status, values_from = count, values_fill = 0) %>%
    mutate(
      Total = overt + none + expletive + `n/a`,
      `Null Rate` = none / (overt + none + expletive)
    )

  kable(neg_wide %>% mutate(`Null Rate` = fmt_pct(`Null Rate`)),
        caption = "Negation position × subject realization",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
}
```

## 5.3 Relative Clauses

Subject vs object relative clause distribution.

```{r relative-clauses}
if (!is.null(DATA[["relative_clauses"]])) {
  rc_data <- train_overall(DATA[["relative_clauses"]])

  if ("metric" %in% names(rc_data) && "value" %in% names(rc_data)) {
    rc_summary <- rc_data %>%
      filter(!is.na(metric), !is.na(value)) %>%
      select(metric, value) %>%
      mutate(value = as.numeric(value))

    kable(rc_summary,
          caption = "Relative clause statistics",
          format.args = list(big.mark = ",")) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
  }
}
```

---

# Part 6: Cross-Genre Summary {.tabset}

## Master Comparison Table

```{r cross-genre-summary}
if (!is.null(DATA[["clause_structure"]]) && !is.null(DATA[["pronoun_inventory"]])) {
  # Null rate by genre (ROOT clauses)
  null_rates <- train_genre(DATA[["clause_structure"]]) %>%
    filter(clause_dep == "ROOT") %>%
    order_genres() %>%
    compute_null_rate(grouping_vars = "genre") %>%
    select(genre, null_rate, expletive_rate)

  # 1st person pronoun rate by genre
  fn_col <- if ("function." %in% names(DATA[["pronoun_inventory"]])) "function." else "function"

  person_rates <- train_genre(DATA[["pronoun_inventory"]]) %>%
    order_genres() %>%
    filter(!is.na(person), person %in% c(1, 2, 3)) %>%
    mutate(person = as.integer(person)) %>%
    group_by(genre, person) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    group_by(genre) %>%
    mutate(pct = count / sum(count)) %>%
    filter(person == 1) %>%
    select(genre, first_person_rate = pct)

  # Combine
  master_summary <- null_rates %>%
    left_join(person_rates, by = "genre") %>%
    mutate(
      modality = ifelse(genre %in% SPOKEN_GENRES, "Spoken", "Written")
    ) %>%
    arrange(modality, genre)

  kable(master_summary %>%
          mutate(
            null_rate = fmt_pct(null_rate),
            expletive_rate = fmt_pct(expletive_rate),
            first_person_rate = fmt_pct(first_person_rate)
          ) %>%
          rename(
            Genre = genre,
            Modality = modality,
            `Null Rate (ROOT)` = null_rate,
            `Expletive Rate` = expletive_rate,
            `1st Person Rate` = first_person_rate
          ),
        caption = "Key rates by genre") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(master_summary, "cross_genre_summary")
}
```

## Genre Clustering

```{r genre-clustering}
if (exists("master_summary") && nrow(master_summary) >= 3) {
  # Prepare data for visualization
  cluster_data <- master_summary %>%
    select(genre, null_rate, first_person_rate) %>%
    mutate(
      modality = ifelse(genre %in% SPOKEN_GENRES, "Spoken", "Written")
    )

  p_cluster <- ggplot(cluster_data,
                      aes(x = first_person_rate, y = null_rate,
                          color = modality, label = genre)) +
    geom_point(size = 4) +
    geom_text(vjust = -1, hjust = 0.5, size = 3) +
    scale_color_manual(values = c("Spoken" = "#2E86AB", "Written" = "#A23B72")) +
    scale_x_continuous(labels = percent_format()) +
    scale_y_continuous(labels = percent_format()) +
    labs(
      title = "Genre Clustering: Null Rate vs 1st Person Rate",
      subtitle = "Spoken genres cluster together (high 1st person, higher null rate)",
      x = "1st Person Pronoun Rate",
      y = "Null Subject Rate (ROOT)",
      color = "Modality"
    ) +
    theme_corpus() +
    theme(axis.text.x = element_text(angle = 0))

  print(p_cluster)
  save_figure(p_cluster, "genre_clustering")
}
```

---

# Part 7: CHILDES Child vs Adult {.tabset}

```{r childes-available-check, results='asis'}
if (!SPEAKER_SPLIT_AVAILABLE) {
  cat('<div class="alert alert-warning">\n')
  cat('<strong>Speaker-split data not yet available.</strong>\n')
  cat('The CHILDES child vs adult split requires a pipeline re-run with speaker annotation.\n')
  cat('This section will populate automatically once the data is available.\n')
  cat('</div>\n')
}
```

## Predictions (Acquisition Literature)

Based on language acquisition research, we predict:

| Phenomenon | Child | Adult | Rationale |
|------------|-------|-------|-----------|
| **Null subjects** | Higher rate | Lower rate | Optional Infinitive stage; early grammar allows subject drop |
| **Non-finite roots** | Higher rate | Lower rate | Root infinitives in early child language |
| **1st person** | Higher rate | Lower rate | Egocentric speech; "I want", "me do it" |
| **2nd person** | Similar | Similar | Both speakers use "you" in dialogue |
| **3rd person** | Lower rate | Higher rate | Adults narrate more; children act |
| **Pre-verbal negation** | Higher rate | Lower rate | "No want", "not go" before restructuring |

```{r childes-analysis, eval=SPEAKER_SPLIT_AVAILABLE}
# Helper to extract child vs adult data
childes_pair <- function(df) {
  df %>%
    filter(split == "train_90M",
           genre %in% c("CHILDES_child", "CHILDES_adult")) %>%
    mutate(role = ifelse(genre == "CHILDES_child", "Child", "Adult"))
}
```

```{r fig-childes-null, eval=SPEAKER_SPLIT_AVAILABLE, fig.width=8, fig.height=5}
if (!is.null(DATA[["clause_structure"]])) {
  cs_ch <- childes_pair(DATA[["clause_structure"]]) %>%
    filter(clause_dep == "ROOT", !is.na(subject_status)) %>%
    compute_null_rate(grouping_vars = "role")

  if (nrow(cs_ch) > 0) {
    p_ch_null <- ggplot(cs_ch %>%
                          pivot_longer(cols = c(overt_rate, null_rate, expletive_rate),
                                       names_to = "status", values_to = "rate") %>%
                          mutate(status = factor(status,
                                                 levels = c("overt_rate", "expletive_rate", "null_rate"),
                                                 labels = c("Overt", "Expletive", "Null"))),
                        aes(x = role, y = rate, fill = status)) +
      geom_col(position = "stack", width = 0.5) +
      scale_fill_manual(values = c("Overt" = "#2E86AB", "Expletive" = "#E8C547", "Null" = "#A23B72")) +
      scale_y_continuous(labels = percent_format()) +
      labs(
        title = "Subject Realization at ROOT: Child vs Adult (CHILDES)",
        x = NULL,
        y = "Proportion",
        fill = "Status"
      ) +
      theme_corpus() +
      theme(axis.text.x = element_text(angle = 0))

    print(p_ch_null)
    save_figure(p_ch_null, "childes_null_comparison")
  }
}
```

```{r fig-childes-person, eval=SPEAKER_SPLIT_AVAILABLE, fig.width=8, fig.height=5}
if (!is.null(DATA[["pronoun_inventory"]])) {
  fn_col <- if ("function." %in% names(DATA[["pronoun_inventory"]])) "function." else "function"

  pi_ch <- childes_pair(DATA[["pronoun_inventory"]]) %>%
    filter(!is.na(person), person %in% c(1, 2, 3)) %>%
    mutate(person = as.integer(person)) %>%
    group_by(role, person) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    group_by(role) %>%
    mutate(pct = count / sum(count)) %>%
    ungroup()

  if (nrow(pi_ch) > 0) {
    p_ch_person <- ggplot(pi_ch, aes(x = factor(person), y = pct, fill = role)) +
      geom_col(position = "dodge", width = 0.6) +
      scale_fill_manual(values = c("Child" = "#17becf", "Adult" = "#e377c2")) +
      scale_y_continuous(labels = percent_format()) +
      labs(
        title = "Pronoun Person Distribution: Child vs Adult (CHILDES)",
        x = "Person",
        y = "Proportion",
        fill = "Speaker"
      ) +
      theme_corpus() +
      theme(axis.text.x = element_text(angle = 0))

    print(p_ch_person)
    save_figure(p_ch_person, "childes_person_comparison")
  }
}
```

---

# Appendix: Reference Tables {.tabset}

## All Clause Structure Data

```{r appendix-clause}
if (!is.null(DATA[["clause_structure"]])) {
  cs_full <- train_overall(DATA[["clause_structure"]]) %>%
    filter(!is.na(clause_dep), clause_dep != "",
           !is.na(subject_status), subject_status != "") %>%
    group_by(clause_dep, subject_status) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = subject_status, values_from = count, values_fill = 0) %>%
    mutate(
      Total = overt + none + expletive,
      `Null %` = fmt_pct(none / Total),
      `Overt %` = fmt_pct(overt / Total)
    ) %>%
    arrange(desc(Total))

  kable(cs_full,
        caption = "Complete clause structure data (train_90M overall)",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(cs_full, "appendix_clause_structure")
}
```

## Top 50 Matrix Verbs

```{r appendix-verbs}
if (exists("verb_data")) {
  top50 <- verb_data %>%
    head(50) %>%
    mutate(Rank = row_number()) %>%
    select(Rank, Verb = verb_lemma, Class = verb_class, Count = count)

  kable(top50,
        caption = "Top 50 matrix verbs for infinitival complements",
        format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(top50, "appendix_matrix_verbs")
}
```

## Top 100 Pronouns

```{r appendix-pronouns}
if (!is.null(DATA[["pronoun_inventory"]])) {
  top100_pronouns <- pi_overall %>%
    filter(!is.na(lemma), lemma != "") %>%
    group_by(lemma) %>%
    summarise(count = sum(as.numeric(count), na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(count)) %>%
    head(100) %>%
    mutate(
      Rank = row_number(),
      pct = count / total_pronouns * 100
    ) %>%
    select(Rank, Lemma = lemma, Count = count, `% of Pronouns` = pct)

  kable(top100_pronouns,
        caption = "Top 100 pronoun lemmas",
        format.args = list(big.mark = ","), digits = 2) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  save_table(top100_pronouns, "appendix_pronouns")
}
```

---

# Summary Statistics

```{r final-summary}
summary_stats <- data.frame(
  Analyzer = ANALYZER_NAMES,
  stringsAsFactors = FALSE
)

get_total <- function(analyzer) {
  df <- DATA[[analyzer]]
  if (is.null(df)) return(NA)
  df_ov <- df %>% filter(split == "train_90M", genre == "overall")

  if ("count" %in% names(df_ov)) {
    return(sum(as.numeric(df_ov$count), na.rm = TRUE))
  } else if ("value" %in% names(df_ov)) {
    total_row <- df_ov %>% filter(metric == "total")
    if (nrow(total_row) > 0) return(as.numeric(total_row$value[1]))
  }
  NA
}

summary_stats$`Total Count` <- sapply(ANALYZER_NAMES, get_total)
summary_stats$Available <- !is.na(summary_stats$`Total Count`)

kable(summary_stats %>%
        mutate(`Total Count` = ifelse(is.na(`Total Count`), "—", fmt_count(`Total Count`))),
      caption = "Analyzer totals (train_90M)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

*Report generated on `r format(Sys.time(), "%Y-%m-%d %H:%M:%S")`*

*Figures saved to `analysis/output/corpus_descriptives/figures/`*

*Tables saved to `analysis/output/corpus_descriptives/tables/`*
